Turbulence and Particle Acceleration in Giant Radio Halos: the Origin of Seed Electrons

### 2016-04-25 paper.tex
some edits in Sect 2

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|{{dp}\over{dt}}\right|_{\rm C} 
+ \frac{p}{3}\left(\bnabla\bcdot \bvel\right)
+ \left|{{dp}\over{dt}}\right|_{\rm r}\right.\right.
\nonumber\\
&-& \left.\left.{1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\bnabla\bcdot \bvel\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum $p$ and time $t$ (which is
normalized such that the number density is given by
$n_{\rmn{e}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/dt=\partial/\partial
t+\bvel\bcdot\bnabla$ is the Lagrangian derivative, $\bvel$ is the gas
velocity, $|dp/dt|$ represents radiative (r) and Coulomb (C)
losses, $D_{pp}$ is the momentum space diffusion coefficient, and
$Q_{\rmn{e}}$ denotes the injection rate of primary and secondary
electrons in the ICM. The first term containing the expression
$\bnabla\bcdot \bvel$ represents Fermi-I acceleration and the second
term of this form describes adiabatic gains and losses.

@@

The time evolution of the spectral energy distribution of CRps,
$f_{\rmn{p}}$, is similarly given by:
\begin{eqnarray}
\lefteqn{
  {{d f_{\rmn{p}}(p,t)}\over{d t}} =
  {{\partial }\over{\partial p}}
  \left\{
  f_{\rmn{p}}(p,t)\left[ \left|{{dp}\over{dt}}\right|_{\rm C}
    + \frac{p}{3}\bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right)\right.\right.}
\nonumber\\
&&-\left.\left. {1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right)
\right]\right\} - f_{\rmn{p}}(p,t) \bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right) 
- \bvel_{\rmn{st}}\bcdot\bnabla f_{\rmn{p}}(p,t)
\nonumber\\
&&+\,\,\, {{\partial^2 }\over{\partial p^2}}
\left[ D_{pp} f_{\rmn{p}}(p,t) \right] - {{f_{\rmn{p}}(p,t)}\over{\tau_{\rm had}(p)}}
+ Q_{\rmn{p}}(p,t)\, ,
\label{eq:FP_p}
\end{eqnarray}
where $\bvel_{\rmn{st}}=-\varv_{\rmn{A}}\bnabla f_{\rmn{p}}/|\bnabla
f_{\rmn{p}}|$ is the streaming velocity in the isotropic transport
approximation, $\varv_{\rmn{A}}$ is the Alfv\'en speed,
$Q_{\rmn{p}}(p,t)$ denotes the injection rate of shock accelerated
CRps as a function of momentum $p$ and time $t$, and $\tau_{\rm had}$
is the timescale of hadronic losses that produce pions via CRp
collisions with thermal protons of the ICM \cite[e.g.][]{brunetti11}.

@@

\begin{table}
  \caption{Thermal quantities and timescales for different spatial
    regions in a Coma like cluster.}
\begin{tabular}{l c  c c}
\hline
\hline
& & spatial regions & \\
 & $0.1\,\RH^{(2)}$ & $0.3\,\RH^{(2)}$ & $\RH^{(2)}$   \\
\hline
thermal quantities$^{(1)}$ & & & \\
$\rho$ [$10^{-27}$ g cm$^{-3}$] & 3.0 & 1.6 & 0.15 \\
$T$ [$10^{8}$ K] & 1.4 & 1.0 & 0.58 \\
\hline
timescales$^{(1)}$ & & & \\
$\tau_\rmn{D}$(\Mprimary)$^{(3)}$ [Gyr] & 0.45 & 0.44 & 0.39 \\
$\tau_\rmn{D}$(\Mstream)$^{(3)}$  [Gyr] & 0.50 & 0.47 & 0.34 \\
$\tau_\rmn{D}$(\Mflatturb)$^{(3)}$  [Gyr] & 0.69 & 0.56 & 0.27 \\
$\tau_\rmn{IC/sync}(P=10^4\,m_ec)^{(4)}$ [Gyr] & 0.11 & 0.15 & 0.22 \\
$\tau_\rmn{had}(P=100\,m_pc)^{(5)}$ [Gyr] & 2.4 & 4.5 & 47 \\
$\tau_\rmn{coul}(P=\,m_ec)^{(6)}$  [Gyr] & 0.0092 & 0.017 & 0.17 \\
\hline
\end{tabular}
\begin{quote}
 Notes: \\ 
 (1) Median quantities from our simulated post-merging cluster g72a derived 
 during last 300~Myrs in time. \\
 (2) Radius of the giant radio halo in Coma where $\RH\approx0.6\,R_{200}$.\\
 (3) Fermi-II reacceleration for both electrons and protons at all energies.\\
 (4) Inverse Compton and synchrotron cooling for electrons.\\
 (5) Catastrophic losses for protons.\\
 (6) Coulomb cooling for electrons (protons factor $m_e/m_p$ smaller).\\
 \label{tab:timescales}
  \end{quote}
\end{table}

@@

We adopt both the density \citep{1992A&A...259L..31B} and temperature
profiles \citep{2009ApJ...696.1886B,2001A&A...365L..67A} derived from
X-ray observations of the Coma cluster. They are given by:
\begin{eqnarray}
n_e &=& \frac{3.4\times10^{-3}}{\left[1+\left(R/R_{\rmn{c}}\right)^2\right]^{1.125}}\nonumber\\
k_{\rmn{B}} T &=& 8.25\,\rmn{keV} \left[1+\left(2R/R_{200}\right)^2\right]^{-0.32}\,.
\end{eqnarray}
The virial and core radii of Coma are given by $R_{200} =
2300\,\rmn{kpc}$ and $R_{\rmn{c}} = 294\,\rmn{kpc}$, respectively.
The bulk of the CRps are injected by relatively low Mach number shocks
and parameterized by $f_\rmn{CRp,inj}(p) =
C_\rmn{CRp,inj}\,p^{-2.5}$. The CRps trace the thermal gas with
$C_\rmn{CRp,inj} \propto \epsilon_\rmn{th}^{\alpha_\rmn{CR,spat}}$,
where the normalization is fixed by assuming that the injected CR
energy in the last 650~Myrs make up 0.03\% of the thermal energy
inside the virial radius, i.e.
$\int_0^{R_{200}}\epsilon_\rmn{CRp,inj} \rmn{d}
V\left(\int_0^{R_{200}}\epsilon_\rmn{th} \rmn{d}V\right)^{-1} =
0.0003$. The spectrum of the initial CRp distribution is determined by
the steady state between injection and cooling,
\begin{equation}
 f_\rmn{CRp,0} \propto \frac{\int_p^\infty f_\rmn{CRp,inj}(p') 
\rmn{d}p'}{\displaystyle\left|{{\rmn{d}p}\over{\rmn{d}t}}\right|_{\rm coul}+\frac{p}{\tau_{\rm had}}}\,,
\end{equation}
where we fix the normalization by requiring
$\int_0^{R_{200}}\epsilon_\rmn{CRp,0}
\rmn{d}V\left(\int_0^{R_{200}}\epsilon_\rmn{th} \rmn{d}V\right)^{-1} =
0.003$. Similarly, the initial CRe distribution is given by the steady
state between cooling (Coulomb, inverse Compton, and synchrotron) and
injection of secondary CRes from $f_{CRp,0}$. We also include the
continuous injection of secondary CRes during $\tau_\rmn{cl}$. The
compressible turbulence responsible for reaccelerating both the CRps
and CRes is parameterized by $I_0 \propto X_\rmn{tu} (n_e
T)^{\alpha_\rmn{tu}}$. The diffusion constant $D_\rmn{pp}$ is
calculated for each radial bin using
Eqns.~\ref{eq:dpp}-\ref{eq:Gamma_e}. All parameters and assumptions
are similar to what is used for our simulated cluster.

@@

We use a simple test-particle model for the CR acceleration and
injection, where each shock inject CRs that trace a power-law in
momentum,
\begin{equation}
  f(p,t) = C(t)\,p^{\alpha_\rmn{inj}}\,,\quad
  \alpha_\rmn{inj}=\frac{(\gamma_\rmn{ad}+1)\mathcal{M}^2}
        {(\gamma_\rmn{ad}-1)\mathcal{M}^2+2}
\end{equation}
determined by the normalization $C(t)$ and the spectral index
$\alpha_{\rmn{inj}}$ that depends on the adiabatic index
$\gamma_\rmn{ad} = 5/3$ and the Mach number of the shock
$\mathcal{M}$. It is given by the ratio of the upstream velocity
($u_2$) and the sound speed ($c_s$). The CR number density and CR
energy density are derived from $n_\rmn{CR} =
\int_{p_\rmn{inj}}^\infty \dd p\, f(p)$ and $\epsilon_\rmn{CR} =
\int_{p_\rmn{inj}}^\infty \dd p\, f(p) \,T(p)$, respectively, where
$T(p) = (\sqrt{1+p^2} -1)\, m\,c^2$ is the kinetic energy of a
particle with momentum $p$. We adopt a fit to Monte Carlo simulations
of the thermal leakage process that relates the momentum of injected
protons ($p_\rmn{inj}$) to the thermal energy ($p_\rmn{th}$) of the
shocked plasma \citep{kang11}:
\begin{eqnarray}
  \label{eq:qinj}
  p_\rmn{inj} &=& x_\rmn{inj} p_\rmn{th} = x_\rmn{inj} \sqrt{\frac{2 \,k T_2}{m c^2}}\,, \nonumber \\
  \rmn{where}\quad x_\rmn{inj} &\approx& 1.17 \frac{u_2}{p_{\rmn th}\,c} \left(1+
  \frac{1.07}{\epsilon_B}\right) \left(\frac{\mathcal{M}}{3}\right)^{0.1}\,.
\end{eqnarray}
Here $\eb = B_0/B_{\perp}$, $B_0$ is the amplitude of the downstream
MHD wave turbulence, and $B_{\perp}$ is the magnetic field along the
shock normal. The physical range of $\eb$ is quite uncertain due to
complex plasma interactions. In this paper, we adopt $\eb = 0.23$,
which we will later see corresponds to a conservative maximum energy
acceleration efficiency for protons of $10\%$. To derive the
acceleration efficiency, $\zeta_\rmn{inj}$, we first have to infer the
particle injection efficiency, which is the fraction of downstream
thermal gas particles which experience diffusive shock acceleration
\citep[for details see ][]{pinzke13},
\begin{equation}
  \label{eq:eta}
  \eta_\rmn{CR,lin} =
  \frac{4}{\sqrt{\upi}}\,\frac{x_\rmn{inj}^3}{\alpha_\rmn{inj}-1}\,
  \rmn{e}^{-x_\rmn{inj}^2}.
\end{equation}
The particle injection efficiency is a strong function of
$x_\rmn{inj}$ that depends on both $\mathcal{M}$ and $\eb$. The
energy density of CRs that are injected and accelerated at the shock
(neglecting the CR back reaction on the shock) is given by
\begin{equation}
\label{eq:CR_energy} 
  \Delta\epsilon_\rmn{CR,lin} =
  \eta_\rmn{CR,lin}(\mathcal{M})\,T_\rmn{CR}(\mathcal{M},p_\rmn{inj})\,n_\rmn{th}(T_2),
\end{equation}
and the CR energy injection and acceleration efficiency is:
\begin{equation}
  \zeta_\rmn{lin} =
  \frac{\Delta\epsilon_\rmn{CR,lin}}{\Delta\epsilon_\rmn{diss}},
   \quad\mbox{where}\quad
  \Delta\epsilon_\rmn{diss} = \epsilon_\rmn{th2} - \epsilon_\rmn{th0}\,\left(\frac{\rho_2}{\rho_0}\right)^{\gamma_\rmn{ad}}\,.
\label{eqn:energy_frac}  
\end{equation}
The dissipated energy density in the downstream regime,
$\Delta\epsilon_\rmn{diss}$, is given by the difference of the thermal
energy densities in the pre- and post-shock regimes, corrected for the
adiabatic energy increase due to gas compression.

>>>>

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|{{dp}\over{dt}}\right|_{\rm Coul} 
+ \frac{p}{3}\left(\bnabla\bcdot \bvel\right)
+ \left|{{dp}\over{dt}}\right|_{\rm rad}\right.\right.
\nonumber\\
&-& \left.\left.{1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\bnabla\bcdot \bvel\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum $p$ and time $t$ (which is
normalized such that the number density is given by
$n_{\rmn{e}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/dt=\partial/\partial
t+\bvel\bcdot\bnabla$ is the Lagrangian derivative, $\bvel$ is the gas
velocity, $|dp/dt|$ represents radiative (rad) and Coulomb (Coul)
losses, $D_{pp}$ is the momentum space diffusion coefficient, and
$Q_{\rmn{e}}$ denotes the injection rate of primary and secondary
electrons in the ICM. The first term containing the expression
$\bnabla\bcdot \bvel$ represents Fermi-I acceleration and the second
term of this form describes adiabatic gains and losses.

@@

The time evolution of the spectral energy distribution of CRps,
$f_{\rmn{p}}$, is similarly given by:
\begin{eqnarray}
\lefteqn{
  {{d f_{\rmn{p}}(p,t)}\over{d t}} =
  {{\partial }\over{\partial p}}
  \left\{
  f_{\rmn{p}}(p,t)\left[ \left|{{dp}\over{dt}}\right|_{\rm Coul}
    + \frac{p}{3}\bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right)\right.\right.}
\nonumber\\
&&-\left.\left. {1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right)
\right]\right\} - f_{\rmn{p}}(p,t) \bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right) 
- \bvel_{\rmn{st}}\bcdot\bnabla f_{\rmn{p}}(p,t)
\nonumber\\
&&+\,\,\, {{\partial^2 }\over{\partial p^2}}
\left[ D_{pp} f_{\rmn{p}}(p,t) \right] - {{f_{\rmn{p}}(p,t)}\over{\tau_{\rm had}(p)}}
+ Q_{\rmn{p}}(p,t)\, ,
\label{eq:FP_p}
\end{eqnarray}
where $\bvel_{\rmn{st}}=-\varv_{\rmn{A}}\bnabla f_{\rmn{p}}/|\bnabla
f_{\rmn{p}}|$ is the streaming velocity in the isotropic transport
approximation, $\varv_{\rmn{A}}$ is the Alfv\'en speed,
$Q_{\rmn{p}}(p,t)$ denotes the injection rate of shock accelerated
CRps as a function of momentum $p$ and time $t$, and $\tau_{\rm had}$
is the timescale of hadronic losses that produce pions via CRp
collisions with thermal protons of the ICM \cite[e.g.][]{brunetti11}.

@@

\begin{table}
  \caption{Thermal quantities and timescales for different spatial
    regions in a Coma like cluster.}
\begin{tabular}{l c  c c}
\hline
\hline
& & spatial regions & \\
 & $0.1\,\RH^{(2)}$ & $0.3\,\RH^{(2)}$ & $\RH^{(2)}$   \\
\hline
thermal quantities$^{(1)}$ & & & \\
$\rho$ [$10^{-27}$ g cm$^{-3}$] & 3.0 & 1.6 & 0.15 \\
$T$ [$10^{8}$ K] & 1.4 & 1.0 & 0.58 \\
\hline
timescales$^{(1)}$ & & & \\
$\tau_\rmn{D}$(\Mprimary)$^{(3)}$ [Gyr] & 0.45 & 0.44 & 0.39 \\
$\tau_\rmn{D}$(\Mstream)$^{(3)}$  [Gyr] & 0.50 & 0.47 & 0.34 \\
$\tau_\rmn{D}$(\Mflatturb)$^{(3)}$  [Gyr] & 0.69 & 0.56 & 0.27 \\
$\tau_\rmn{IC/sync}(P=10^4\,m_ec)^{(4)}$ [Gyr] & 0.11 & 0.15 & 0.22 \\
$\tau_\rmn{had}(P=100\,m_pc)^{(5)}$ [Gyr] & 2.4 & 4.5 & 47 \\
$\tau_\rmn{Coul}(P=\,m_ec)^{(6)}$  [Gyr] & 0.0092 & 0.017 & 0.17 \\
\hline
\end{tabular}
\begin{quote}
 Notes: \\ 
 (1) Median quantities from our simulated post-merging cluster g72a derived 
 during last 300~Myrs in time. \\
 (2) Radius of the giant radio halo in Coma where $\RH\approx0.6\,R_{200}$.\\
 (3) Fermi-II reacceleration for both electrons and protons at all energies.\\
 (4) Inverse Compton and synchrotron cooling for electrons.\\
 (5) Catastrophic losses for protons.\\
 (6) Coulomb cooling for electrons (protons factor $m_e/m_p$ smaller).\\
 \label{tab:timescales}
  \end{quote}
\end{table}

@@

We adopt both the density \citep{1992A&A...259L..31B} and temperature
profiles \citep{2009ApJ...696.1886B,2001A&A...365L..67A} derived from
X-ray observations of the Coma cluster. They are given by:
\begin{eqnarray}
n_e &=& \frac{3.4\times10^{-3}}{\left[1+\left(R/R_{\rmn{c}}\right)^2\right]^{1.125}}\nonumber\\
k_{\rmn{B}} T &=& 8.25\,\rmn{keV} \left[1+\left(2R/R_{200}\right)^2\right]^{-0.32}\,.
\end{eqnarray}
The virial and core radii of Coma are given by $R_{200} =
2300\,\rmn{kpc}$ and $R_{\rmn{c}} = 294\,\rmn{kpc}$, respectively.
The bulk of the CRps are injected by relatively low Mach number shocks
and parameterized by $f_\rmn{CRp,inj}(p) =
C_\rmn{CRp,inj}\,p^{-2.5}$. The CRps trace the thermal gas with
$C_\rmn{CRp,inj} \propto \epsilon_\rmn{th}^{\alpha_\rmn{CR,spat}}$,
where the normalization is fixed by assuming that the injected CR
energy in the last 650~Myrs make up 0.03\% of the thermal energy
inside the virial radius, i.e.
$\int_0^{R_{200}}\epsilon_\rmn{CRp,inj} \rmn{d}
V\left(\int_0^{R_{200}}\epsilon_\rmn{th} \rmn{d}V\right)^{-1} =
0.0003$. The spectrum of the initial CRp distribution is determined by
the steady state between injection and cooling,
\begin{equation}
 f_\rmn{CRp,0} \propto \frac{\int_p^\infty f_\rmn{CRp,inj}(p') 
\rmn{d}p'}{\displaystyle\left|{{\rmn{d}p}\over{\rmn{d}t}}\right|_{\rm Coul}+\frac{p}{\tau_{\rm had}}}\,,
\end{equation}
where we fix the normalization by requiring
$\int_0^{R_{200}}\epsilon_\rmn{CRp,0}
\rmn{d}V\left(\int_0^{R_{200}}\epsilon_\rmn{th} \rmn{d}V\right)^{-1} =
0.003$. Similarly, the initial CRe distribution is given by the steady
state between cooling (Coulomb, inverse Compton, and synchrotron) and
injection of secondary CRes from $f_{CRp,0}$. We also include the
continuous injection of secondary CRes during $\tau_\rmn{cl}$. The
compressible turbulence responsible for reaccelerating both the CRps
and CRes is parameterized by $I_0 \propto X_\rmn{tu} (n_e
T)^{\alpha_\rmn{tu}}$. The diffusion constant $D_\rmn{pp}$ is
calculated for each radial bin using
Eqns.~\ref{eq:dpp}-\ref{eq:Gamma_e}. All parameters and assumptions
are similar to what is used for our simulated cluster.

@@

We use a simple test-particle model for the CR acceleration and
injection, where each shock inject CRs that trace a power-law in
momentum,
\begin{equation}
  f(p,t) = C(t)\,p^{\alpha_\rmn{inj}}\,,\quad
  \alpha_\rmn{inj}=\frac{(\gamma_\rmn{ad}+1)\mathcal{M}^2}
        {(\gamma_\rmn{ad}-1)\mathcal{M}^2+2}
\end{equation}
determined by the normalization $C(t)$ and the spectral index
$\alpha_{\rmn{inj}}$ that depends on the adiabatic index
$\gamma_\rmn{ad} = 5/3$ and the Mach number of the shock
$\mathcal{M}$. It is given by the ratio of the upstream velocity
($u_2$) and the sound speed ($c_s$). The CR number density and CR
energy density are derived from $n_\rmn{CR} =
\int_{p_\rmn{inj}}^\infty \dd p\, f(p)$ and $\epsilon_\rmn{CR} =
\int_{p_\rmn{inj}}^\infty \dd p\, f(p) \,T(p)$, respectively, where
$T(p) = (\sqrt{1+p^2} -1)\, m\,c^2$ is the kinetic energy of a
particle with momentum $p$. We adopt a fit to Monte Carlo simulations
of the thermal leakage process that relates the momentum of injected
protons ($p_\rmn{inj}$) to the thermal energy ($p_\rmn{th}$) of the
shocked plasma \citep{kang11}:
\begin{eqnarray}
  \label{eq:qinj}
  p_\rmn{inj} &=& x_\rmn{inj} p_\rmn{th} = x_\rmn{inj} \sqrt{\frac{2 \,k T_2}{m c^2}}\,, \nonumber \\
  \rmn{where}\quad x_\rmn{inj} &\approx& 1.17 \frac{u_2}{p_{\rmn{th}}\,c} \left(1+
  \frac{1.07}{\epsilon_B}\right) \left(\frac{\mathcal{M}}{3}\right)^{0.1}\,.
\end{eqnarray}
Here $\eb = B_0/B_{\perp}$, $B_0$ is the amplitude of the downstream
MHD wave turbulence, and $B_{\perp}$ is the magnetic field along the
shock normal. The physical range of $\eb$ is quite uncertain due to
complex plasma interactions. In this paper, we adopt $\eb = 0.23$,
which we will later see corresponds to a conservative maximum energy
acceleration efficiency for protons of $10\%$. To derive the
acceleration efficiency, $\zeta_\rmn{inj}$, we first have to infer the
particle injection efficiency, which is the fraction of downstream
thermal gas particles which experience diffusive shock acceleration
\citep[for details see ][]{pinzke13},
\begin{equation}
  \label{eq:eta}
  \eta_\rmn{CR,lin} =
  \frac{4}{\sqrt{\upi}}\,\frac{x_\rmn{inj}^3}{\alpha_\rmn{inj}-1}\,
  \rmn{e}^{-x_\rmn{inj}^2}.
\end{equation}
The particle injection efficiency is a strong function of
$x_\rmn{inj}$ that depends on both $\mathcal{M}$ and $\eb$. The
energy density of CRs that are injected and accelerated at the shock
(neglecting the CR back reaction on the shock) is given by
\begin{equation}
\label{eq:CR_energy} 
  \Delta\epsilon_\rmn{CR,lin} =
  \eta_\rmn{CR,lin}(\mathcal{M})\,T_\rmn{CR}(\mathcal{M},p_\rmn{inj})\,n_\rmn{th}(T_2),
\end{equation}
and the CR energy injection and acceleration efficiency is:
\begin{equation}
  \zeta_\rmn{lin} =
  \frac{\Delta\epsilon_\rmn{CR,lin}}{\Delta\epsilon_\rmn{diss}},
   \quad\mbox{where}\quad
  \Delta\epsilon_\rmn{diss} = \epsilon_\rmn{th2} - \epsilon_\rmn{th0}\,\left(\frac{\rho_2}{\rho_0}\right)^{\gamma_\rmn{ad}}\,.
\label{eqn:energy_frac}  
\end{equation}
The dissipated energy density in the downstream regime,
$\Delta\epsilon_\rmn{diss}$, is given by the difference of the thermal
energy densities in the pre- and post-shock regimes, corrected for the
adiabatic energy increase due to gas compression.
